---
title: "NFL Formation Analysis"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
knitr::opts_chunk$set(error = FALSE)
#setwd('~/OMSA/CDA/Final Project/Data')
```

Import libraries.

```{r}
library(jtools)
library(car)
library(glmnet)
```

Import and format csv file created with NFL_defense_clustering.ipynb

```{r}
dat = read.csv('r_2_file.csv', stringsAsFactors = TRUE)
dat$cluster_=as.factor(dat$cluster_)
dat$Personnel = as.factor(dat$Personnel)
print(head(dat))
```

Run ANOVA on EPA by formation clusters to determine if there is a significant difference in offensive play success against different clusters.

```{r}
anov1= aov(epa~cluster_, data=dat)
summary(anov1)
```

Perform a Tukey Pairwise analysis of EPA by cluster.

```{r}
TukeyHSD(anov1, 'cluster_', conf.level=0.99)
```

Run ANOVA again with interaction terms.

```{r}
anov2 = lm(epa~cluster_+play_type_+cluster_:play_type_, data=dat)
summ(anov2)
```

Examine assumptions of ANOVA to determine if there are any serious violations.

```{r}
par(mfrow=c(2,2))
plot(anov2$fitted.values, anov2$residuals, xlab = 'residuals', ylab='fitted values', main = 'Residuals vs. Fitted')
qqnorm(anov2$residuals)
qqline(anov2$residuals)
hist(anov2$residuals, main = '', xlab='residuals')
```
2-way ANOVA with interaction terms included.



Apply Box-Cox transformation to EPA.

```{r}
dat$epa = dat$epa-min(dat$epa)+1
model2  = lm(epa~cluster_+play_type_+cluster_:play_type_, data=dat)
bc = boxCox(model2)
opt.lambda = bc$x[which.max(bc$y)]
# Round it to the nearest 0.5
cat("Optimal lambda:", opt.lambda, end="\n")
```
```{r}
anov3 = lm(epa**opt.lambda~cluster_+play_type_+cluster_:play_type_, data=dat)
summ(anov3)

par(mfrow=c(2,2))
plot(anov3$fitted.values, anov3$residuals, xlab = 'residuals', ylab='fitted values', main = 'Residuals vs. Fitted')
qqnorm(anov3$residuals)
qqline(anov3$residuals)
hist(anov3$residuals, main = '', xlab='residuals')
```

Run ridge regression model with all variables

```{r}
library(glmnet)
X =  model.matrix( ~ .-1, dat[,-10])
ridge.cv = cv.glmnet(X, dat[,10], alpha=1, nfolds = 30)
ridge.model <- glmnet(X, dat[,10], alpha=1, nlambda = 100)
coef(ridge.model, ridge.cv$lambda.min)
```
Forward-backward stepwise regression.

```{r}
n = nrow(dat)
base_mod = lm(epa~1, data=dat)
full_mod = lm(epa~.+cluster_:play_type_, data=dat)
stepwise_model = step(base_mod, scope=list(lower=base_mod, upper=full_mod), direction="both", k=2, trace=F)
summ(stepwise_model)
```

```{r}
step(base_mod, scope=list(lower=base_mod, upper=full_mod), direction="both", k=2)
```